### 2022-11-26 manuscript/figures_tables.Rmd

---
output: 
  bookdown::word_document2:
    reference_docx: word_styles_reference_01.docx
    fig_caption: true
    number_sections: false
bibliography: [citations.bib]
csl: apa.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
knitr::opts_knit$set(root.dir = "/Users/bjpessman/Documents/phd_research_code/Agelenopsis_aggregation/manuscript")
require(knitr)
```

```{r import, eval = FALSE, include = FALSE}
webs <- read.table(file = "data/webs.txt", header = TRUE) 
webs <- webs %>% 
  mutate(ID = factor(ID),
         Location = factor(Location), 
         Location = fct_recode(Location, 
                               "Urban Forest" = "Forest", 
                               "Urban Center" = "Campus"),
         Location = fct_relevel(Location, "Urban Forest", "Urban Center"),
         Land = fct_recode(Land, 
                           "Urban, High" = "UrbanizedHighIntensity",
                           "Urban, Medium" = "UrbanizedMediumIntensity",
                           "Urban, Low" = "UrbanizedLowIntensity",
                           "Deciduous Forest" = "DeciduousForest",
                           "Woody Wetlands" = "WoodyWetlands"),
         Land = fct_relevel(Land, 
                            "Urban, High", 
                            "Urban, Medium", 
                            "Urban, Low", 
                            "Deciduous Forest", 
                            "Woody Wetlands"),
                  patch_area_km = patch_area_mm * 1e-6) %>% 
  dplyr::rename("Impervious" = "Imperv", 
                "Dist_to_Road" = "road_dist_data", 
                "Traffic_Dist_Road" = "dist_traffic_road", 
                "Dist_to_Highway" = "highway_dist_m", 
                "Traffic_Dist_Highway" = "dist_traffic_highway") #data on individual webs

sites <- webs %>% 
  filter(Web == "W001")

web_dist <- webs %>% 
  mutate(Location = factor(Location)) %>% 
  filter(RetreatDist != 0,
         RetreatDist < 1000,
         Web == "W002" | Web == "W003") %>% 
  mutate(Neighbor = ifelse(Web == "W002", "1", "2"))

near <- webs %>% 
  mutate(Location = factor(Location)) %>% 
  filter(RetreatDist != 0,
         RetreatDist < 1000,
         Occ_Neigh == "1")

substrates <- read.table(file = "data/substrate.txt", header = TRUE)

site_info <- read.table(file = "data/site_info.txt", header = TRUE)

land_cover <- read.table(file = "data/land_cover_table.txt", header = TRUE) 
```

```{r spider, dpi = 300}
knitr::include_graphics("/Users/bjpessman/Documents/phd_research_code/Agelenopsis_aggregation/manuscript/figures/spider.png")
```

**Fig. 1**

\newpage

```{r land_location, fig.width = 6.85, dpi = 300}
knitr::include_graphics("/Users/bjpessman/Documents/phd_research_code/Agelenopsis_aggregation/manuscript/figures/land_locations.png")
```

**Fig. 2**

\newpage

```{r main, fig.width = 6.85, fig.height = 7, dpi = 300}
main <- readRDS("figures/main.Rds")
main
```

**Fig. 3**

\newpage

```{r pred_abund, fig.width = 6.85, fig.height = 7, dpi = 300}
predictors_abund <- readRDS("figures/predictors_abund.Rds")
predictors_abund
```

**Fig. 4**

\newpage

```{r pred_agg, fig.width = 6.85, fig.height = 7, dpi = 300}
predictors_agg <- readRDS("figures/predictors_agg.Rds")
predictors_agg
```

**Fig. 5**

\newpage

## Tables

**Table 1**

```{r landscape, fig.width = 3.77, dpi = 300, out.width="50%"}
knitr::include_graphics("/Users/bjpessman/Documents/phd_research_code/Agelenopsis_aggregation/manuscript/figures/landscape.png")
```

\newpage

**Table 2**

```{r pred_location, fig.width = 6.85, dpi = 300}
knitr::include_graphics("/Users/bjpessman/Documents/phd_research_code/Agelenopsis_aggregation/manuscript/figures/pred_location.png")
```

\newpage

**Table 3**

```{r pred_table, fig.width = 6.85, dpi = 300}
knitr::include_graphics("/Users/bjpessman/Documents/phd_research_code/Agelenopsis_aggregation/manuscript/figures/pred_table.png")
```