### 2024-09-30 5_environmental_aic_selection.Rmd

---
title: "Environmental predictors of spider spatial patterns "
author: "Brandi Pessman"
date: "`r Sys.Date()`"
output: html_document
---
# Setting the Working Directory and Global Code Chunk Options

I have set message and warning to FALSE for every chunk so the knitted document does not include messages and warnings.

```{r setup}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, echo = FALSE)
require("knitr")
opts_knit$set(root.dir = "/Users/bjpessman/Documents/phd_research_code/Agelenopsis_aggregation")
```

# Packages to Load

```{r packages}
library(tidyverse) # for making graphs
library(pander) # viewing AIC comparisons between models
library(AICcmodavg) # also helps viewing AIC comparisons
library(MuMIn) # to run the dredge function
library(lme4) # to run the glmer function for the web distance (random effects)
source('functions/AIC_function.R') # runs AIC model selection
```

# Import the Data from the wrangled_data/ folder

```{r import}
sites_wrangled <- readRDS('wrangled_data/sites_wrangled.RDS')
neighbors_wrangled <- readRDS('wrangled_data/neighbors_wrangled.RDS')
webs_wrangled <- readRDS('wrangled_data/webs_wrangled.RDS')
```

# Not Used: One method of model selection

```{r unused}
# This is one way to look at the overall data, but ended up not using this information
models = list(WalkDist ~ 1,
              WalkDist ~ log_TotalSub, 
              WalkDist ~ log_tdr, 
              WalkDist ~ log_tdh,
              WalkDist ~ log_TotalSub + log_tdr,
              WalkDist ~ log_TotalSub + log_tdh,
              WalkDist ~ log_tdr + log_tdh,
              WalkDist ~ log_TotalSub + log_tdr + log_tdh)
fits = lapply(models, glm, data = sites_wrangled, family = "poisson")
modnames = sapply(models, function(ff)deparse(ff[[3]]))
pander(aictab(fits, modname = modnames), caption="AICc model selection table for Search Distance to the Focal Web Overall.", split.tables = Inf)
# The top model keeps all 3 variables and no other model is within 2 Model Likelihood Units
summary(glm(WalkDist ~ log_TotalSub + log_tdr + log_tdh, data = sites_wrangled, family = poisson))
```

# Results

```{r results}
aic_selection(y_var = "NumWebs", data = sites_wrangled)
aic_selection(y_var = "NumSpiders", data = sites_wrangled)
aic_selection(y_var = "round(RetreatDist, 0)", data = neighbors_wrangled)

# web height
#overall
global <- glmer.nb(round(RetreatHeight, 0) ~ log_TotalSub + log_tdr + log_tdh + (1 | ID), data = webs_wrangled)
options(na.action = "na.fail") # Required for dredge to run
model_dredge <- dredge(global, beta = "none", evaluate = TRUE, rank = AICc)
options(na.action = "na.omit") # set back to default
top_model <- get.models(model_dredge, subset = 1)[[1]]
summary(top_model)

r.squaredGLMM(top_model) # trigamma R2m
model.sel(model_dredge) #estimates same signs
summary(model.avg(model_dredge, subset = delta <= 2))
car::vif(get.models(model_dredge, 5)[[1]]) # looking for less than 2

# Urban Forest
webs_forest <- webs_wrangled %>% 
  filter(Location == "Urban Forest") %>% 
  mutate(tree100m_frac = I(tree100m/100))
global <- glmer.nb(round(RetreatHeight, 0) ~ tree100m_frac + log_TotalSub + log_tdr + log_tdh + log_patch + (1 | ID), data = webs_forest, control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e4)))
options(na.action = "na.fail") # Required for dredge to run
model_dredge <- dredge(global, beta = "none", evaluate = TRUE, rank = AICc)
options(na.action = "na.omit") # set back to default
top_model <- get.models(model_dredge, subset = 1)[[1]]
summary(top_model)

r.squaredGLMM(top_model) #trigamma R2m
model.sel(model_dredge) #estimates same signs
summary(model.avg(model_dredge, subset = delta <= 2))
#car::vif(get.models(model_dredge, 1)[[1]]) # looking for less than 2

# Urban Center
webs_center <- webs_wrangled %>% 
  filter(Location == "Urban Center")
webs_center$road_length_m_sc <- scale(webs_center$road_length_m, center = FALSE)
global <- glmer.nb(round(RetreatHeight, 0) ~ log_TotalSub + log_tdr + log_tdh + scale(spec_rad, center = FALSE) + log_patch + road_length_m_sc + (1 | ID), data = webs_center)#, control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e4)))
options(na.action = "na.fail") # Required for dredge to run
model_dredge <- dredge(global, beta = "none", evaluate = T, rank = AICc)
options(na.action = "na.omit") # set back to default
top_model <- get.models(model_dredge, subset = 1)[[1]]
summary(top_model)

r.squaredGLMM(top_model) 
model.sel(model_dredge) #estimates same signs
summary(model.avg(model_dredge, subset = delta <= 2))
#car::vif(get.models(model_dredge, 1)[[1]]) # looking for less than 2
```