### 2024-09-30 6_land_cover_posthoc_nb.Rmd

---
title: "Land cover effects on web abundance, aggregation, and position"
author: "Brandi Pessman"
date: "`r Sys.Date()`"
output: html_document
---
# Setting the Working Directory and Global Code Chunk Options

I have set message and warning to FALSE for every chunk so the knitted document does not include messages and warnings.

```{r setup}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
require("knitr")
opts_knit$set(root.dir = "/Users/bjpessman/Documents/phd_research_code/Agelenopsis_aggregation")
```

# Packages to Load

```{r packages}
library(MASS) # negative binomial tests
library(multcomp) # for pairwise comparisons of land cover categories
library(lme4) # to run the glmer function for the web distance (random effects)
```

# Import the Data from the wrangled_data/ folder

```{r import}
sites_wrangled <- readRDS('wrangled_data/sites_wrangled.RDS')
neighbors_wrangled <- readRDS('wrangled_data/neighbors_wrangled.RDS')
webs_wrangled <- readRDS('wrangled_data/webs_wrangled.RDS')
```

# Results of the negative binomial and post hoc analyses

```{r results}
# search distance
search_land.nb <- glm.nb(WalkDist ~ Land, data = sites_wrangled)
summary(search_land.nb)
car::Anova(search_land.nb, test.statistic = "LR")
comp <- glht(search_land.nb, linfct = mcp(Land = "Tukey"))
summary(comp)

# number of webs
webs_land_nb <- glm.nb(NumWebs ~ Land, data = sites_wrangled)
summary(webs_land_nb)
car::Anova(webs_land_nb, test.statistic = "LR")
comp <- glht(webs_land_nb, linfct = mcp(Land = "Tukey"))
summary(comp)

# number of spiders
spiders_land_nb <- glm.nb(NumSpiders ~ Land, data = sites_wrangled)
summary(spiders_land_nb)
car::Anova(spiders_land_nb, test.statistic = "LR")
comp <- glht(spiders_land_nb, linfct = mcp(Land = "Tukey"))
summary(comp)

# neighbor distances
dist_land_nb <- glm.nb(round(RetreatDist, 0) ~ Land, data = neighbors_wrangled)
summary(dist_land_nb)
car::Anova(dist_land_nb, test.statistic = "LR")
comp <- glht(dist_land_nb, linfct = mcp(Land = "Tukey"))
summary(comp)

# web height
height_land_nb <- glmer.nb(round(RetreatHeight, 0) ~ Land + (1 | ID), data = webs_wrangled)
summary(height_land_nb)
car::Anova(height_land_nb, type = 3)
comp <- glht(height_land_nb, linfct = mcp(Land = "Tukey"))
summary(comp)
```